---
title: "baseballr Presentation"
author: "Ian Curtis"
date: "2022-11-06"
output: pdf_document
---

# Load Packages

I first load the necessary packages.

```{r load_packages, message = FALSE}
library(baseballr)
library(tidyverse)
```

# Introduction

`baseballr` is a package used to collect data on various baseball (MLB) statistics from multiple sources on the internet. It can also provide some interesting data on pre-selected trends and certain calculations.

As of October 2022, `baseballr` is capable of fetching data from the following sources:

* The MLB API
* The MLB Statcast database (Baseball Savant)
* Retrosheet
* NCAA
* Baseball Reference
* FanGraphs

This package is incredibly useful for searching for MLB data, especially when dataset joining is needed (such as combining statistics from Baseball Reference and FanGraph, for instance).

# Demonstration

Below are a few ways in which the `baseballr` package might be used to grab data.

## Statcast (Baseball Savant)

The [Savant database](https://baseballsavant.mlb.com/) is a large, searchable repository of MLB data extending back to 2008. The database can be searched on the web and contains a large number of custom filters to apply. The database will automatically create aggregate summaries according to selections, but the raw data is pitch-by-pitch and gives the researcher much freedom when using the data.

The package here will grab raw data based on the query which can either specify a specific batter or pitcher or request all of the raw data between a certain time frame.

```{r load_scherzer_data, message=FALSE}
# Search for all data for Max Scherzer in June 2021

scherzer <- statcast_search_pitchers(
  start_date = '2021-06-01', 
  end_date = '2021-06-30', 
  pitcherid = 453286)
head(scherzer)
```

The above chunk searches for all of the pitch-by-pitch data for Max Scherzer in June 2021. The result is a large data frame containing attributes that can be pulled out for use.

From here, we might plot some data!

```{r plot_scherzer}
scherzer_plot <- scherzer %>% 
  ggplot(aes(x = release_speed, y = release_spin_rate, color = pitch_type)) +
  geom_point() +
  labs(title = 'Max Scherzer: Release Speed vs. Ball Spin Rate', 
       subtitle = 'Broken down by pitch type', 
       x = 'Release Speed (MPH)', 
       y = 'Release Spin Rate (RPM)') +
  guides(color = guide_legend(title = "Pitch Type")) +
  scale_color_brewer(palette = "Dark2")

scherzer_plot
```

## Baseball Reference

Baseball Reference is another source of baseball data. The package `baseballr` allows for aggregate player performance data to be scraped as well as historical standings at any date. There is also a function to calculate "team consistency". Baseball Reference might be used more for getting "typical" statistics such as batting average, ERA, and number of home runs.

```{r load_bref_data, message=F}
bref_batter <- bref_daily_batter("2021-06-01", "2021-06-30")

head(bref_batter)
```

```{r plot_bref}
bref_batter %>% 
  filter(PA >= 15) %>% 
  ggplot(aes(x = BA, y = OBP)) +
  geom_point() +
  labs(title = 'Batting Average vs. On-Base Percentage for Batters in June 2021', 
       subtitle = 'Minimum 15 Plate Appearances')
```


# References

* https://baseballsavant.mlb.com/
* https://razzball.com/mlbamids/
* https://www.baseball-reference.com/
* https://billpetti.github.io/baseballr/index.html

